---
title: "Bootstrap_Practical"
format: html
---

### Bootstrapping

##### You only have one sample, how do you get a confidence interval?

You make use of bootstrapping. Using your own bootstraps to solve your issue.

You only have one sample in BS world, make the assumption that that sample is representative of the population.

Resample from that sample. Treat original sample as the population.

The \* represents the bootstrap. The BS distribution is for the distribution (spread) around something, like original sample estimate.

##### How to sample

With replacement: Probability with picking random values from a population.

Without replacement: just get same sample back.

##### Why use BS?

Std error

Bias

Conf. Interval

BS is the way to understand behaviour of an estimator.

##### Bootstrap Principle

Treat sample (\^) as if its the population, then sample from that (\*)

Assume \* relative to \^ is the same as \^ relative to population. But \^ is cenetred at observed statistics.

***Bootstrap distr to estimate sampling distribution.***

*Look st slides for Bias, SE and CI.*

##### Basic and Percentile BS Intervals

2 different things, specify which one you've used.

look at notes for the formula

use R quantiles() to get the values for the CI.

Another problem in BS: is when we have skew data. How to get around that: Use BS T-dist to then get the CI etc.

```{r}
#?sample()
set.seed(1)
unif <- runif(50, 0, 99)
u <- sample(unif, size = 10, replace = TRUE)
u_without <- sample(u, size = 10, replace = FALSE)

B <- 10
bstr <- matrix(0,ncol = 50,nrow =B)
for(i in 1:B){
  samp = sample(u,size = 50,replace =TRUE)
  bstr[i,]  = samp
}
```

#### Question 1

```{r}
library(MASS)   # Modern Applied Statistics with S 
                # (Venables and Ripley, 2004)

gal <- galaxies     # data set on velocities (km/s) 
                    #of 82 galaxies

hist(galaxies, breaks = 25)
```

```{r}
## MEDIAN speed of galaxies
tobs <- median(gal)       # T(X)
```

```{r}
set.seed(1)

B <- 10000
bstr <- matrix(0,ncol = 82,nrow =B)
for(i in 1:B){
  samp = sample(gal,size = 82,replace =TRUE)
  bstr[i,]  = samp
}

tboot = apply(bstr,1,median)
```

```{r}
#mean of medians - median(sample)
bias <- mean(tboot) - tobs
bias
```

```{r}
SE_tboot <- sd(tboot)
SE_tboot 
```

It decreases the estimate of the standard error of your distribution.

```{r}
#Plot the BS samples, and include tobs 
hist(tboot, breaks = 25) 
abline(v= c(tobs, mean(tboot)), col= c("orange", "pink"), lwd = 2)
```

```{r}
c1 <- quantile(tboot, probs = (0.05/2))
c2 <- quantile(tboot, probs = 1-0.05/2)
#For the medians
#Percentile CI
perc_CI <- c(c1, c2) #no need to sort as quantile does this 

#Basic BS CI
CI_basic <- c(2*tobs - c2, 2*tobs - c1)
names(CI_basic) <- c("2.5%", "97.5%")
print(perc_CI)
print(CI_basic)
```

```{r}
tobs_mean <- mean(gal)
tboot_mean <- apply(bstr, 1, mean)

c1 <- quantile(tboot_mean, probs = (0.05/2))
c2 <- quantile(tboot_mean, probs = 1-0.05/2)

#For the means: Percentile 
mperc <- c(c1, c2)
print(mperc)

#Means: Basic Method 
mbasic_CI <- c(2*tobs_mean - c2, 2*tobs_mean - c1)
names(mbasic_CI) <- c("2.5%", "97.5%")
print(mbasic_CI)

#t_CI
n <- length(tboot_mean)
se <- sd(tboot_mean)/sqrt(n)

t_star <- (tboot_mean-tobs_mean)/se

c1 <- quantile(t_star, probs = (0.05/2))
c2 <- quantile(t_star, probs = 1-0.05/2)

CI_t <- c(tobs_mean - c2*se, tobs_mean - c1*se) 
names(CI_t) <- c("2.5%", "97.5%")
CI_t
```

#### Question 2

Why is it that my values differ? Is there an error in my calculations??

```{r}
library(boot)

bt.smpls <- boot(gal, function(x, i) median(x[i]), R = 3000) #simulate 3000 BS replicates        

# see Venables and Ripley, pg.134
# take 3000 bootstrap samples, returns the 3000 medians of these

bt.smpls
summary(bt.smpls)

boot.ci(bt.smpls, type = c("norm", "basic", "perc", "bca"))     # all sorts of confidence intervals
help(boot.ci)           # bca gives a bias corrected and accelerated 
                       # (improved) percentile interval
```

#### Question 3

```{r}
# Filter to only the columns we need and remove rows with NAs
clean_data <- na.omit(airquality[, c("Ozone", "Temp")])

# Now the function is super simple
corr_func <- function(data, indices) {
  d <- data[indices, ]
  return(stats::cor(d$Ozone, d$Temp))
}

# Run it
boot_results <- boot(data = clean_data, statistic = corr_func, R = 10000)
boot.ci(boot_results, type = c("perc", "bca"))
```

```{r}
#Find std error, bias and CI for cor
#Std Error 
set.seed(1)
boot(data = airquality, statistic = corr_func, R =10000)
```

```{r}
n <- dim(df)[1]

# sample cases from data frame
cases <- sample(1:n, replace = T)
booti <- df[cases, ]               
```
